\documentclass[10pt,letterpaper]{article}
\usepackage[margin=.75in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{pgfplots}
\usepackage[shortlabels]{enumitem}
\usepackage{listings}
\usepackage[document]{ragged2e}
\usepackage{inconsolata}
\pgfplotsset{compat=1.16}
\usepackage{graphicx}
\usepackage{setspace}
\graphicspath{ {./images/} }
\onehalfspacing

\author{Gabriel Brown}
\title{Statistics For Data Science: HW 1}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\lstset{frame=tb,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}
\setlength{\headheight}{25pt}
\fancyhf{}
\rhead{
    Rajan Puri | Shameem Ahmed Khan | Orkun Akyol | Gabriel Brown\\
    Statistics | Winter 2024\\
    Homework 1
}
\rfoot{\thepage}

\begin{document}
\paragraph{1}
\textbf{A medical test for disease D has outcomes + (positive) and - (negative). We assume that \begin{itemize}
    \item the probability for an individual to have the disease is 0.01
    \item the probability of a positive test, given that the individual has the disease, is 0.9
    \item the probability of a negative test, given that the individual does not have the disease, is 0.9
\end{itemize}
Compute the probability that an individual has the disease, given that the
individual has tested positive. Comment on the quality of the test.}\\
P(has disease $|$ positive test) = P(positive test $|$ has disease) P(has disease) / P(positive test)\\
= $0.9 * 0.01 /$ P(positive test)\\
P(positive test) = P(positive test $|$ has disease) P(has disease) + P(positive test $|$ doesn't have disease) P(doesn't have disease)\\
= P(positive test $|$ has disease) P(has disease) + (1 - P(negative test $|$ doesn't have disease)) P(doesn't have disease)\\
= $0.9 * 0.01 + (1 - 0.9)*.99 = 0.108$\\
Therefore, the probability that a patient has the disease given a positive test = $0.9 * 0.01 / 0.108 = 0.0833$ or 8.33\%\\
Therefore, the test is not very useful in diagnosing a patient.

\paragraph{2}
\textbf{The Monty Hall problem is a probability puzzle based on a game show scenario:\begin{itemize}
    \item There are three doors: behind one door is a prize (e.g., a car) and behind the other two doors are goats.
    \item A contestant chooses one of the three doors.
    \item The host, who knows what is behind each door, opens one of the two
remaining doors with a goat behind it.
    \item The contestant is then given the option to either stick with their initial choice or switch to the other unopened door.
\end{itemize}
Your task is to write a program (using Python or any other programming
language of your choice) to simulate the probability of winning the prize if
the contestant always switches their initial choice after the host reveals a goat.
You can model the scenario as follows:\begin{itemize}
    \item The prize is randomly assigned to one of the three doors (labeled 1,2 or 3) in a uniformly random manner.
    \item The contestant randomly selects one of the doors (1, 2 or 3) with equal probability.
    \item One of the unchosen doors that does not have the prize behind it is
eliminated.
    \item The contestant switches their initial choice to the remaining unopened door.
\end{itemize}
Simulate the above process $10^5$ times and track the number of times the contestant wins the prize by switching their initial choice. Based on this data,
compute the proportion of times the contestant wins using this strategy. In
general, do you have a better chance at winning the prize by changing your
initial choice?
}\\
See code in problem2.py\\
Output:\\
\begin{lstlisting}[language=Python]
total wins: 66719
win percentage when switching: 66.719
\end{lstlisting}
If we don't change our initial choice, our chance of winning is 1 in 3, or 33.3\%\\
Experimentally, we've shown that if we always switch, our change of winning is much higher than this.\\
Therefore, based on this data, you have a better chance at winning in general if you always switch.

\paragraph{3}
\textbf{Let X be a continuous real-valued random variable with the probability density
function $f: \mathbb{R} \xrightarrow{} \mathbb{R}$,
$$f(x) = \begin{cases}
    0 & \text{ if } x < 0\\
    \frac{4xe^{-x^2}}{(1+e^{-x^2})^2} & \text{ if } x \geq 0
\end{cases}$$}\\
\begin{enumerate}[(a)]
    \item \textbf{Compute the cumulative distribution function $F(x) = \mathbb{P}(X \leq x)$ of $X$ for $x \in \mathbb{R}$}\\
    For all $x < 0$, $F(x) = 0$ trivially\\
    For all $x \geq 0, F(x) = \int_{0}^{x} f(u) du$\\
    $= \int_{0}^{x} \frac{4ue^{-u^2}}{(1+e^{-u^2})^2} du$\\
    $= \frac{-2}{e^{u^2}+1}|_{0}^{x} = \frac{e^{x^2}-1}{e^{x^2}+1} = \tanh{\frac{x^2}{2}}$\\
    Therefore,
    $F(x) = \begin{cases}
        0 & x < 0\\
        \tanh(\frac{x^2}{2}) & x \geq 0\\
    \end{cases}$

    \item \textbf{Solve the quantile function of X}\\
    Given $q \in (0,1), \exists x$ such that $F(x) = q$\\
    $\implies \tanh(\frac{x^2}{2}) = q, \forall q\in (0,1)$\\
    $\implies x = \sqrt{2\tanh^{-1}(q)}$\\
    $\implies F_X^{-1}(q)= \sqrt{2\tanh^{-1}(q)}$ for $q\in (0,1)$

    \item \textbf{Compute the probability $\mathbb{P}(0 < X < 1)$. Which value $a\in \mathbb{R}$ satisfies $\mathbb{P}(X \leq a) = 0.95$}\\
    $P(0 < X < 1) = F(1) - F(0) = \tanh(1/2) \approx 0.462$\\
    by the definition of the quantile function, $a = F_X^{-1}(.95) = \sqrt{2\tanh^{-1}(.95)} \approx 1.91404$
\end{enumerate}


\paragraph{4}
\begin{enumerate}[(a)]
    \item The sample space is the following set with four members: 
        \[
            \omega = \{(0,0), (1,0), (0,1), (1,1)\}
        \]
        By simple enumeration, it follows:\\ 
        \text{} \\
    $p_X(x) = \begin{cases}
        1/4 & \text{if } x = 0\\
        1/2 & \text{if } x = 1\\
        1/4 & \text{if } x = 2\\
    \end{cases}$
    \item Trivially, $F(x)=0$ for $x<0$ and $F(x)=1$ for $x \ge 2$. By cumulatively summing the probabilities in the probability mass function, we get: \\
    \text{} \\
    $F(x) = \begin{cases}
        0 & \text{if } x<0\\
        1/4 & \text{if } 0 \le x < 1\\
        3/4 & \text{if } 1 \le x < 2\\
        1 & \text{if } x \ge 2\\
    \end{cases}$
    \item $F^{-1}(q) = \begin{cases}
         0 & \text{if } 0 < q \le 1/4\\
         1 & \text{if } 1/4 < q \le 3/4\\
         2 & \text{if } 3/4 < q < 1\\
    \end{cases}$
\end{enumerate}
\end{document}
